Molecular alignment at graphene surface determines tansport properties of graphene transistors

### 2020-04-01 manuscript.tex
some corrections
finished introduction and added refs to .bib file.

  Graphene field-effect transistor structures were used to investigate the role of molecular alignment on charge transport properties of heterostructures comprising a single-layer graphene and variable thickness of N,N'-bis(n-octyl)-(1,7\&1,6)-dicyanoperylene-3,4:9,10-bisdicarboximide (PDI8-CN2) - an n-type organic semiconductor. Our atomic force microscopy data show that under selected growth conditions PDI8-CN2 grows in a layer-by-layer fashion up to a second monolayer. The firs layer comprises flat-lying molecules, whereas the molecules in the second layer orient themselves in an upright orientation. Transconductance measurements show that the flat-lying molecules have little effect on the position of the Fermi level in graphene. Upright oriented molecules in the second layer instead have a strong effect as to neutralize native $p$-type doping of graphene and cause a shift of charge-neutrality level towards the Dirac point. We interpret such behavior in terms of different orientation of the surface dipole on layers with different molecular orientations. At the same time the overall mobility of the charge carriers reaches values exceeding 3000 cm$^{2}$/Vs. We 
\end{abstract}

@@

One of the venue to improve switching capabilities of graphene in potential high-speed electronic applications\cite{novoselov-2004} includes coupling with organic semiconductors (OS). In this regard  blending of graphene nanoflakes with OS\cite{pathipati-2019,pathipati-2015,pathipati-2014} or templated growth of OS on graphene\cite{zhou-2013a,nouchi-2008,nouchi-2014}. Charge transport in blends can be described in terms of combination of hopping through the manifold of localized states on OS molecules and extended-state transport on graphene. Less clear is the situation in the case of OS overlayers on graphene. As we show in this work, charge carrier transport in such material systems is crucially dependent on the molecular orientation at the graphene/OS interface.

For this type of research it is important to select OS that exhibits high charge carrier mobility per se, in order to preserve high mobility on graphene.   N,N'-bis(n-octyl)-(1,7\&1,6)-dicyanoperylene-3,4:9,10-bisdicarboximide (PDI8-CN2)
PDI8-CN2 exhibits a remarkable air stability, solution processability and relatively high electron mobility in thin film form\cite{jung-2014,molinari-2009,piliego-2009a}.  The position of the lowest unoccupied molecular orbital (LUMO) in this material is at 7.1 eV\cite{jung-2014}, which makes it suitable for interfacing with graphene, whose Dirac point is at 4.4 eV \textbf{*Vadym: check*}

OS growth on graphene is characterized by...

@@

We observe transfer characteristics of GFET comprising clean graphene exhibiting charge-neutrality point at back-gate voltage of 8.5 V.  The position of CNP relative to the Dirac point (0 V) is consistent with p-type doping of graphene caused by the trap states on SiO$_{2}$ surface and chemical redox reaction occurring at the graphene/SiO$_{2}$ interface\cite{feng-2014,peng-2017,lafkioti-2010}.  In addition to the shift of the conductivity minimum we observe also a strong reduction in the slope of the transfer curve occurring at about 3.6 V, leading to a plateau of the drain current.  Such behavior is commonly present in GFETs with relatively short channels, where the role of the metallic contacts becomes important\cite{bartolomeo-2015,nouchi-2008,nouchi-2014}. Two principal mechanisms in GFET affect charge transport under ambient conditions. First is already mentioned doping of the channel with airborne species. The second is related to the graphene doping under metallic contacts\cite{farmer-2009,giovannetti-2008,lee-2008,mueller-2009}.  Several groups report the evidence that metallic contacts induce doping of the underlying graphene \cite{giovannetti-2008,mueller-2009,huard-2008}, and the amount of transferred charge depends on the difference of metallic work function and the work function of graphene.  Specifically, Au contacts used in our experiments cause p-type doping, according to the work by Giovanetti et al.\cite{giovannetti-2008}, causing a shift in Fermi level of 0.3 eV in the direction away from the vacuum level, across the Dirac point. In addition, changes in the interface charge density may be a consequence of Fermi-level depinning at the metal/graphene interface and is related to microscopic interface environment that is characterized by interdiffusion of metallic atoms and/or oxide interlayer\cite{nouchi-2008,nouchi-2014}. Since Fermi level at such interfaces is not pinned or only partially pinned\cite{dibartolomeo-2011}, the interface charge density may be affected by the applied gate voltage, which is not the case for the abrupt interfaces.  In the channel instead, we can calculate the position of the Fermi level relative to the Dirac point in graphene ($\Delta E$) from the measured position of CNP (Figure \ref{fig:3}), using the approach from Ref. \onlinecite{wang-2008}:
\begin{equation}
  \label{eq:1}
  \Delta E = \hbar v_{F}\sqrt{\pi\alpha|V_{g}-V_{D}}
\end{equation}

@@

Firstly, we focus on the initial charge carrier density $n_0$, which stems from the impurities residing either at the graphene interface or within the gate dielectric. On clean graphene we obtained a value of 3.33$\times 10^{11}$ cm$^{-2}$, which may be considered a ``clean'' sample for long-range scattering.  As the coverage of PDI8-CN2 is increased to 0.7 $n_0$ exhibits only a slight increase to 3.71$\times 10^{11}$ cm$^{-2}$, but almost doubles at full coverage ($n_{0}=7.42\times 10^{11}$ cm$^{-2}$).
In order to explain the observed behavior, we have to consider the dependence of the electronic structure of organic thin layers on the molecular orientation.  Anisotropic behavior of charge transport properties of organic thin films are known for some time now\cite{sundar-2004}. Less common are detailed studies of the dependence of the position of the electron energy levels at different molecular orientation. This is particularly true for n-type OSs on graphene. Zhang et al.\cite{zhang-2016d} have performed a systematic photoelectron spectroscopy study coupled to density functional theory (DFT) -based calculations of electrostatic potential surfaces of several organic semiconductors including n-type. Their findings demonstrate energy shifts of the electronic structures as high as 0.54 eV between the layers with the molecules oriented upright and flat-lying molecules.  Crucial for such difference is the existence of the dipole (and quadrupole) electric field arising at different moieties of the molecules. For example, in the case of pyrrolo[3,4-c]pyrrole-1,4-dione (DPP)  the calculations show a surface electric dipole due to the electron-withdrawing carbonyl and imine groups pointing  away from the surface when the molecules are oriented upright (shorter molecular axis perpendicular to the surface). On the other hand flat-lying DPP molecules exhibit a quadrupole moment that points towards the surface. Consequently, the energy of HOMO is closer to the vacuum level in the case of flat-lying molecules relative to upright oriented molecules.

>>>>

  Graphene field-effect transistor structures were used to investigate the role of molecular alignment on charge transport properties of heterostructures comprising a single-layer graphene and variable thickness of N,N'-bis(n-octyl)-(1,7\&1,6)-dicyanoperylene-3,4:9,10-bisdicarboximide (PDI8-CN2) - an n-type organic semiconductor. Our atomic force microscopy data show that under selected growth conditions PDI8-CN2 grows in a layer-by-layer fashion up to a second monolayer. The firs layer comprises flat-lying molecules, whereas the molecules in the second layer orient themselves in an upright orientation. Transconductance measurements show that the flat-lying molecules have little effect on the position of the Fermi level in graphene. Upright oriented molecules in the second layer instead have a strong effect as to neutralize native $p$-type doping of graphene and cause a shift of charge-neutrality level towards the Dirac point. We interpret such behavior in terms of different orientation of the surface dipole on layers with different molecular orientations. At the same time the overall mobility of the charge carriers reaches values exceeding 3000 cm$^{2}$/Vs.  
\end{abstract}

@@

One of the venue to improve switching and optical capabilities of graphene in potential high-speed electronic applications\cite{novoselov-2004} or photodetectors includes coupling with organic semiconductors (OS). In this regard  two basic approaches are most common. Blending of graphene nanoflakes with OS  has been demonstrated to effective both in polymer OS\cite{pathipati-2015,elgemayel-2014,liscio-2011} or small-molecule OS \cite{pathipati-2014}. The other approach includes interfacing graphene with OS overlayers\cite{pathipati-2019,pathipati-2015,huisman-2015,kim-2017b,ha-2012,zhou-2013a,nouchi-2008,nouchi-2014}. Charge transport in blends can be described in terms of combination of hopping through the manifold of localized states on OS molecules and extended-state transport on graphene. Less clear is the situation in the case of OS overlayers on graphene, especially when the OS layer thickness remains in the range of few molecular layers. In this case, the electric field due to the charge imbalance on the molecules may still have an important effect on the position of the Fermi level in graphene. Most of the published experiments including those presented here include mechanically exfoliated graphene flakes transferred onto SiO$_{2}$ in air. The resulting redox reactions at the substrate surface cause $p$-type doping of graphene \cite{feng-2014,peng-2017,lafkioti-2010}, which may be changed upon deposition of OS molecules either via charge transfer or dipole/multipole-induced charge accumulation at the interface. This inevitably alters charge-transporting properties of the final graphene/OS system, and must be considered when interpreting current-voltage ($I-V$) characteristics of graphene field-effect-transistors (GFETs) that include thin OS overlayers.

Research in initial stages of growth of OS on dielectric substrates is plentiful and a thorough review of the subject is beyond the scope of this article. It is important to note, however that initial stages of growth of OS are extremely sensitive to all three relevant parameters: arriving rate of the molecules, their surface mobility and the rate of island nucleation. The work by Ruiz et al. \cite{ruiz-2004}, for example, summarizes rich morphology that can be achieved by altering growth conditions of pentacene on Si$_{2}$. Similar findings hold for other small-molecule OSs as well as polymers grown from the solution. The importance of the ability to control the initial stages of growth can't be overemphasized, since it is in this stage that the morphology of thicker layer is determined, and the coupling between the layer morphology and its charge transport properties are well known.

As for the graphene-based electronic devices initial stages of growth are particularly important since in this stage the orientation of molecules on the substrate are determined. The geometrical orientation of the electric dipoles and multipoles depends, in turn on the orientation of the molecule on the substrate\cite{geng-2012}. The electric field dipole exhibits a $\propto 1/r^{3}$ dependence on the distance from the dipole and is thus important only for the extremely thin overlayes on graphene. In addition it is important that the overlayers are structurally coherent i.e. monocrystalline domains are as large as possible in order to reduce disorder-related reduction in the effects of surface dipoles. Indeed, as  we show in this work, charge carrier transport in such material systems is crucially dependent on the molecular orientation at the graphene/OS interface. For this type of research it is important to select OS that exhibits high charge carrier mobility per se, in order to preserve high mobility on graphene.   N,N'-bis(n-octyl)-(1,7\&1,6)-dicyanoperylene-3,4:9,10-bisdicarboximide (PDI8-CN2)
PDI8-CN2 exhibits a remarkable air stability, solution processability and relatively high electron mobility in thin film form\cite{jung-2014,molinari-2009,piliego-2009a}.  The position of the lowest unoccupied molecular orbital (LUMO) in this material is at 7.1 eV\cite{jung-2014}, which makes it suitable for interfacing with graphene, whose Dirac point is at $~4.5$ eV.

Our results show that under the selected growth conditions PDI8-CN2 forms well ordered layers up to the second layer. The growth mode is rather complex since the molecules are oriented flat on the graphene surface in the first layer. In the second layer the molecules are oriented almost perpendicular to the surface. This has profound consequences on the measured transfer characteristics. Due to the orientation of the surface dipole of the molecules in the first layer its effect on the transport properties of graphene in almost negligible, and transfer characteristics of the clean graphene relative to that of the 70\% covered with PDI8-CN2 are very similar. On the other hand when the second PDI8-CN2 layer is completed its surface dipole causes a major effect on the position of the charge neutrality level in graphene so that it almost coincides with the position of the Dirac point.  The overall mobility of the structure is reduced relative to the clean graphene, albeit reaches values higher than 3400 cm$^{2}$/Vs, which is a remarkable value for the OS/graphene system.

@@

We observe transfer characteristics of GFET comprising clean graphene exhibiting charge-neutrality point at back-gate voltage of 8.5 V.  The position of CNP relative to the Dirac point (0 V) is consistent with $p$-type doping of graphene caused by the trap states on SiO$_{2}$ surface and chemical redox reaction occurring at the graphene/SiO$_{2}$ interface\cite{feng-2014,peng-2017,lafkioti-2010}.  In addition to the shift of the conductivity minimum we observe also a strong reduction in the slope of the transfer curve occurring at about 3.6 V, leading to a plateau of the drain current.  Such behavior is commonly present in GFETs with relatively short channels, where the role of the metallic contacts becomes important\cite{bartolomeo-2015,nouchi-2008,nouchi-2014}. Two principal mechanisms in GFET affect charge transport under ambient conditions. First is already mentioned doping of the channel with airborne species. The second is related to the graphene doping under metallic contacts\cite{farmer-2009,giovannetti-2008,lee-2008,mueller-2009}.  Several groups report the evidence that metallic contacts induce doping of the underlying graphene \cite{giovannetti-2008,mueller-2009,huard-2008}, and the amount of transferred charge depends on the difference of metallic work function and the work function of graphene.  Specifically, Au contacts used in our experiments cause p-type doping, according to the work by Giovanetti et al.\cite{giovannetti-2008}, causing a shift in Fermi level of 0.3 eV in the direction away from the vacuum level, across the Dirac point. In addition, changes in the interface charge density may be a consequence of Fermi-level depinning at the metal/graphene interface and is related to microscopic interface environment that is characterized by interdiffusion of metallic atoms and/or oxide interlayer\cite{nouchi-2008,nouchi-2014}. Since Fermi level at such interfaces is not pinned or only partially pinned\cite{dibartolomeo-2011}, the interface charge density may be affected by the applied gate voltage, which is not the case for the abrupt interfaces.  In the channel instead, we can calculate the position of the Fermi level relative to the Dirac point in graphene ($\Delta E$) from the measured position of CNP (Figure \ref{fig:3}), using the approach from Ref. \onlinecite{wang-2008}:
\begin{equation}
  \label{eq:1}
  \Delta E = \hbar v_{F}\sqrt{\pi\alpha|V_{g}-V_{D}}
\end{equation}

@@

Firstly, we focus on the initial charge carrier density $n_0$, which stems from the impurities residing either at the graphene interface or within the gate dielectric. On clean graphene we obtained a value of 3.33$\times 10^{11}$ cm$^{-2}$, which may be considered a ``clean'' sample for long-range scattering.  As the coverage of PDI8-CN2 is increased to 0.7 $n_0$ exhibits only a slight increase to 3.71$\times 10^{11}$ cm$^{-2}$, but almost doubles at full coverage ($n_{0}=7.42\times 10^{11}$ cm$^{-2}$).
In order to explain the observed behavior, we have to consider the dependence of the electronic structure of organic thin layers on the molecular orientation.  Anisotropic behavior of charge transport properties of organic thin films are known for some time now\cite{sundar-2004}. Less common are detailed studies of the dependence of the position of the electron energy levels at different molecular orientation. This is particularly true for $n$-type OSs on graphene. Zhang et al.\cite{zhang-2016d} have performed a systematic photoelectron spectroscopy study coupled to density functional theory (DFT) -based calculations of electrostatic potential surfaces of several organic semiconductors including n-type. Their findings demonstrate energy shifts of the electronic structures as high as 0.54 eV between the layers with the molecules oriented upright and flat-lying molecules.  Crucial for such difference is the existence of the dipole (and quadrupole) electric field arising at different moieties of the molecules. For example, in the case of pyrrolo[3,4-c]pyrrole-1,4-dione (DPP)  the calculations show a surface electric dipole due to the electron-withdrawing carbonyl and imine groups pointing  away from the surface when the molecules are oriented upright (shorter molecular axis perpendicular to the surface). On the other hand flat-lying DPP molecules exhibit a quadrupole moment that points towards the surface. Consequently, the energy of HOMO is closer to the vacuum level in the case of flat-lying molecules relative to upright oriented molecules.
